---
title: "Null models network metrics"
author: "Carine Emer"
date: "28/08/2017"
output: html_document
---


```{r pakg}
library(bipartite)
library(vegan)
```


```{r input data}
setwd("~/Dropbox/Seed dispersal data Atlantic/Metanetwork_data/MS_Carine/data")
nested_int<-read.csv("nested spp interactions new.csv", header=TRUE, row.names = 1)
nested_test<-read.csv("nested test.csv", header=TRUE, row.names = 1)
nested_birds<-read.csv("nested spp birds new.csv", header=TRUE, row.names = 1)
nested_plants<-read.csv("nested spp plants new.csv", header=TRUE, row.names = 1)
```

From oecosimu
#"swap": sequential algorithm for binary matrices that changes the matrix structure, but does not influence marginal sums (Gotelli & Entsminger 2003). This inspects 2 by 2 submatrices so long that a swap can be done.
#"tswap": sequential algorithm for binary matrices. Same as the "swap" algorithm, but it tries a fixed number of times and performs zero to many swaps at one step (according to the thin argument in the call). This approach was suggested by Miklós & Podani (2004) because they found that ordinary swap may lead to biased sequences, since some columns or rows are more easily swapped.

## quasiswap - recommened by Carsten/Oksanen, is an optimized version of ‘swaps’ that reduces numbers above one to maintain marginal totals

## r1,r2 preserves the site (row) frequencies, take into account commonness and rarity of species


FROM NULLMODELS
##swap.web -- MT, C constant --- biased, and only for quantitative matrices
## Shuffle.web maintain the dimensionality of the network (number of links and interactions) but not the marginal totals, ie. the sum of observed interactions per fragment (rows) and the sum of interactions frequency (columns).



```{r mod}

m_int<-computeModules(nested_int)
listModuleInformation(m_int) ## 16 modules ## ALWAYS THE SAME

mod_birds<-computeModules(nested_birds)
mod_birds ### VARIES A BIT
printoutModuleInformation(mod_birds)
listModuleInformation(mod_birds) # four modules ---- VERY UNSTABLE

mod_plants<-computeModules(nested_plants)
printoutModuleInformation(mod_plants)
listModuleInformation(mod_plants) ## 12 modules

# Give here the mean observed modularity values
mod_obs_int <- 0.738207
mod_obs_b <- 0.3396158
mod_obs_p <- 0.5725531


#########NULL MODELS
### Modularity

#Msig <- function (mat, mlike)  {
    require(bipartite)
    # mat is the input matrix for which M is tested
    # mlike is the observed mean M value
    nulls <- nullmodel(nested_birds, N=100, method=4) ## 2 = swapweb, connectance and marginal totals constant (ie, constrait degree for sites and interactions/spp)
    modules.nulls <- sapply(nulls, computeModules)
    like.nulls <- sapply(modules.nulls, function(x) x@likelihood)
    z <- (mlike - mean(like.nulls))/sd(like.nulls)
    p <- 2*pnorm(-abs(z))}
    #cat("\n\n","P value for modularity M= ", mlike, "\n", 
        "No. resamplings= ", N, "\n",
        "P=  ",p)} 
#m_birds<-Msig(nested_birds, mod_obs_b) ### 
#m_plants<-Msig(nested_plants, mod_obs_p)
#m_int<-Msig(nested_int, mod_obs_int)



```


## which null model to chose? tswap (Miklos & Podani 2004, quasiswap - recommened by Carsten) 

##from vegan:oecosimu
```{r nulls nestedness}
## null models for NESTEDNESS

null_r1<-oecosimu(nested_int,nestednodf,"r2",nsimul = 100, order=FALSE) ## info at commsim {vegan}
densityplot(permustats(null_r1))

# tswap - NODF ns
# quasiswap and r2 - NODF significant

#### BIRDS
null_r1<-oecosimu(nested_birds,nestednodf,"r2",nsimul = 100, order=FALSE) ## info at commsim {vegan}
null_r1
densityplot(permustats(null_r1))

#### PLANTS
null_plants<-oecosimu(nested_plants,nestednodf,"r2",nsimul = 100, order=FALSE) ## info at commsim {vegan}
null_plants
densityplot(permustats(null_r1))

```


## from vegan:nullmodel
```{r nulls connectance}
obs<-networklevel(nested_int, index="connectance")

nulls <- (simulate(vegan::nullmodel(nested_int, method="quasiswap"), nsim = 100))

m1<-nulls[,,1]
m2<-nulls[,,2]
identical(m1,m2)

res.nulls <- apply(nulls, 3, networklevel, index= "connectance")
hist(res.nulls)
z<-obs-mean(res.nulls)/sd(res.nulls)
praw <- sum(res.nulls>(obs)) / length(res.nulls)
ifelse(praw > 0.5, 1-praw, praw)

p <- 2*pnorm(-abs(z))
p

mean(res.nulls)
sd(res.nulls)


#------------------- BIRDS
obs_cB<-networklevel(nested_birds, index="connectance")

nulls <- (simulate(vegan::nullmodel(nested_birds, method="r2"), nsim = 100))
res.nulls <- apply(nulls, 3, networklevel, index= "connectance")
hist(res.nulls)
z<-obs_cB-mean(res.nulls)/sd(res.nulls)
p <- 2*pnorm(-abs(z))
p
mean(res.nulls)
sd(res.nulls)
z
#------------------- PLANTS
obs_cP<-networklevel(nested_plants, index="connectance")

nulls <- (simulate(vegan::nullmodel(nested_plants, method="r2"), nsim = 100))
res.nulls <- apply(nulls, 3, networklevel, index= "connectance")
hist(res.nulls)
z<-obs_cP-mean(res.nulls)/sd(res.nulls)
p <- 2*pnorm(-abs(z))
p
mean(res.nulls)
sd(res.nulls)
z
```

```{r nulls modularity}

######---------------------------- INTERACTIONS
m_int<-computeModules(nested_int)
mod_obs<-0.7379964

nulls_qswap <- (simulate(vegan::nullmodel(nested_int, method="quasiswap"), nsim = 100))
    mod.nulls_qswap <- apply(nulls_qswap, 3, computeModules)
    like.nulls.qswap <- sapply(mod.nulls_qswap, function(x) x@likelihood)
    z_qswap <- (mod_obs - mean(like.nulls.qswap))/sd(like.nulls.qswap) #sd = 0.0008127112
    z_qswap
    p <- 2*pnorm(-abs(z))
    p

#####
nulls_mod2 <- (simulate(vegan::nullmodel(nested_int, method="r2"), nsim = 100))
    mod.nulls2 <- apply(nulls_mod2, 3, computeModules)
    like.nulls2 <- sapply(mod.nulls2, function(x) x@likelihood)
    z2 <- (mod_obs - mean(like.nulls2))/sd(like.nulls)#sd=0.004645199, #sdr1 = 0.005419148
    z2
    p <- 2*pnorm(-abs(z))
    p
    
#####----------------------------- BIRDS
mod_birds<-computeModules(nested_birds) ## 0.3413494
obs_birds<- 0.3413494   
nulls_qswap_b <- (simulate(vegan::nullmodel(nested_birds, method="r2"), nsim = 100))
    mod.nulls_qswap_b <- apply(nulls_qswap_b, 3, computeModules)
    like.nulls.qswap_b <- sapply(mod.nulls_qswap_b, function(x) x@likelihood)
    z_qswap_b <- (obs_birds - mean(like.nulls.qswap_b))/sd(like.nulls.qswap_b) #sd =0.004423643 , mean = 0.3081806, obs = 0.3413494
    z_qswap_b #7.498081
    p_b <- 2*pnorm(-abs(z_qswap_b))
    p_b  #6.475873e-14  
  
    
#### PLANTS    
    nulls_qswap_p <- (simulate(vegan::nullmodel(nested_plants, method="r2"), nsim = 100))
    mod.nulls_qswap_p <- apply(nulls_qswap_p, 3, computeModules)
    like.nulls.qswap_p <- sapply(mod.nulls_qswap_p, function(x) x@likelihood)
    z_qswap_p <- (obs_plants - mean(like.nulls.qswap_p))/sd(like.nulls.qswap_p) #sd = 0.002320598, mean= 0.5596753, obs = 0.571716
    z_qswap_p #5.188602
    p_p<- 2*pnorm(-abs(z_qswap_p))
    p_p #2.118791e-07
    
    
```




## from bipartite:nullmodels - mgen
```{r nulls connectance mgen}

###### CONNECTANCE
obs_c<-networklevel(nested_int, index="connectance")

null1<-nullmodel(nested_int, N=100, method="mgen")
null <- unlist(sapply(null1, networklevel, index="connectance")) 

praw <- sum(null>obs_c) / length(null)
ifelse(praw > 0.5, 1-praw, praw)    # P-value

mean_c_mgen<-mean(null)
sd_c_mgen<-sd(null)

z<-(obs_c-mean_c_mgen)/sd_c_mgen
z

# --------------------- Birds
obs_b<-networklevel(nested_birds, index="connectance")

obs_bnull1<-nullmodel(nested_birds, N=100, method="mgen")
null_b <- unlist(sapply(obs_bnull1, networklevel, index="connectance")) 

praw <- sum(null_b>obs_b) / length(null_b)
ifelse(praw > 0.5, 1-praw, praw)    # P-value

mean_c_mgenb<-mean(null_b)
sd_c_mgenb<-sd(null_b)

z<-(obs_b-mean_c_mgenb)/sd_c_mgenb
z

# --------------------- PLANTS
obs_p<-networklevel(nested_plants, index="connectance")

obs_pnull1<-nullmodel(nested_plants, N=100, method="mgen")
null_p <- unlist(sapply(obs_pnull1, networklevel, index="connectance")) 

praw <- sum(null_p>obs_p) / length(null_p)
ifelse(praw > 0.5, 1-praw, praw)    # P-value

mean_c_mgenp<-mean(null_p)
sd_c_mgenp<-sd(null_p)

z<-(obs_p-mean_c_mgenp)/sd_c_mgenp
z
```

```{r nulls nestedness mgen}
obs_nodf<-nested(nested_int,method= "NODF")
obs_nodfb<-nested(nested_birds,method= "NODF")


################ NESTEDNESS MGEN - EQUIPROBABLE

null_nodf_mgen <-nullmodel(nested_birds, N=100, method="mgen") 
null1_nodf_mgen <-sapply (X=null_nodf_mgen, FUN=nested, method="NODF") ## sapply to aplly the NODF for each random netw
mean_null1_nodf_mgen <- apply (X=cbind(null1_nodf_mgen),MARGIN=2, FUN=mean, na.rm=T) 
sd.null1_nodf_mgen <- apply(X=cbind(null1_nodf_mgen), MARGIN=2, FUN=sd, na.rm=T)
z<-(obs_nodfb-mean_null1_nodf_mgen)/sd.null1_nodf_mgen
z
praw <- sum(null1_nodf_mgen < (obs_nodf)) / length(null1_nodf_mgen)
ifelse(praw > 0.5, 1-praw, praw)

p_nodf_mgen <- 2*pnorm(-abs(z))
p_nodf_mgen

plot(density(null1_nodf_mgen), lwd=2, xlim=c(0,40))
abline(v=obs_nodfb, col="red", lwd=2)

#------------------- PLANTS
obs_nodfp<-nested(nested_plants,method= "NODF")
null_nodf_mgen <-nullmodel(nested_plants, N=100, method="mgen") 
null1_nodf_mgen <-sapply (X=null_nodf_mgen, FUN=nested, method="NODF") 
mean_null1_nodf_mgen <- apply (X=cbind(null1_nodf_mgen),MARGIN=2, FUN=mean, na.rm=T) 
sd.null1_nodf_mgen <- apply(X=cbind(null1_nodf_mgen), MARGIN=2, FUN=sd, na.rm=T)
z<-(obs_nodfp-mean_null1_nodf_mgen)/sd.null1_nodf_mgen
z


```

```{r nulls mod mgen}
## INTERACTIONS mgen - equiprobable
    nulls_mgen <- nullmodel(nested_int, N=100, method=5) 
    mod.nulls_mgen <- sapply(nulls_mgen, computeModules)
    like.nulls_mgen <- sapply(mod.nulls_mgen, function(x) x@likelihood)
    z_mod_mgen <- (0.738 - mean(like.nulls_mgen))/sd(like.nulls_mgen)# sd=0.005392437
    z_mod_mgen
    p <- 2*pnorm(-abs(z_mod_mgen))

##### BIRDS
mod_birds<-computeModules(nested_birds) ## 0.3413494
obs_birds<- 0.3413494   
    nulls_mgenb <- nullmodel(nested_birds, N=100, method=5) 
    mod.nulls_mgenb <- sapply(nulls_mgenb, computeModules)
    like.nulls_mgenb <- sapply(mod.nulls_mgenb, function(x) x@likelihood)
    z_mod_mgenb <- (obs_birds - mean(like.nulls_mgenb))/sd(like.nulls_mgenb)# sd=0.005392437, mean = 0.3720894, z = -4.011063
    z_mod_mgenb
    p <- 2*pnorm(-abs(z_mod_mgenb))    
        
    
    
    
##### PLANTS    
mod_plants<-computeModules(nested_plants) ## 0.3413494
obs_plants<- 0.571716
    nulls_mgenp <- nullmodel(nested_plants, N=100, method=5) 
    mod.nulls_mgenp <- sapply(nulls_mgenp, computeModules)
    like.nulls_mgenp <- sapply(mod.nulls_mgenp, function(x) x@likelihood)
    z_mod_mgenp <- (obs_plants - mean(like.nulls_mgenp))/sd(like.nulls_mgenp)# sd=sd(like.nulls_mgenp), mean = 0.6093741, z = -4.049711
    z_mod_mgenp
    p <- 2*pnorm(-abs(z_mod_mgenp))   
    
    

```








